## Course: Visual Analytics for Policy and Management
###Multivariate Homework 
####Group 3: Dave Coomes, Quinhas Fernndes, Isabella Sun, Long Zong


##
##
```{r echo=T, include=F}
library(uwIntroStats)
library(ggplot2)
library(dplyr)
library(readr)
library(ggiraph)
library(ggiraphExtra)
library(foreign)
library(haven)
library(dotwhisker)
library(broom)
library(dplyr)
library(openxlsx)

knitr::opts_chunk$set(include = T)

link="https://github.com/ihsun-uw/Group3_Final_Project/raw/master/child_mortality.dta"

df <- read_dta(url(link))
```

Create a dataframe with the relevant variabels to plot in our radar plot
```{r}
# copy data
df1=df

df_aggre1 <- aggregate(cbind(birthatend, hwdensity, midwifedensity, mddensity)~provname, data=df1, FUN=mean )

df_aggre1 <- df_aggre1[-c(5), ]

# get minimun value by row
df_aggre1$min=apply(df_aggre1[,c(2:5)],1,min)

# turn this min values into a ranking
df_aggre1$min=rank(df_aggre1$min,ties.method ='first' )

# order city by ranking
prov_fact=as.factor(df_aggre1[order(df_aggre1$min),]$provname)

# turn city into ordered factor
df_aggre1$provname=factor(df_aggre1$provname,
                   levels= prov_fact,
                   labels = prov_fact,
                   ordered = T)

# delete column with ranks
df_aggre1$min=NULL

colnames(df_aggre1) <- c("provname", "Health Professional Present at Birth", "Health Worker Density", "Midwife Density", "Medical Doctor Density")

```

Create Plot
```{r fig.width=15, fig.height=10,echo=TRUE}
base = ggRadar(df_aggre1,aes(group='provname'),legend.position="none") 

radar1 = base + facet_wrap(~provname,nrow =2) +
    labs(title = "Radar plot: Province main study variables",
                    caption = "Fig.1: Describes how provinces are performing on study variables : Health professional present at birth, health worker density, midwife density, and medical doctor density. 

Source:Health Information System") +
              theme(panel.background = element_rect(fill = "gray90"),
                    plot.caption = element_text(hjust = 0, size = 10), 
                    plot.title = element_text(hjust = 0.5, size=14)) 

radar1 
```

We want to see the possible factors that may contribute to the declined mortality rate in Mozambique, and which one is more effective than others.

```{r warning=FALSE}
scat44 <- ggplot(df, aes()) + 
  geom_point(aes(x=year, y=hwdensity, color="Health Worker Density"), size = 0.8, alpha=1/3) + 
  geom_smooth(aes(x=year, y=hwdensity), method="loess", se=T, size=1, color="purple") +
  geom_point(aes(x=year, y=midwifedensity, color= "Midwife Density"), size = 0.8, alpha=1/4) + 
  geom_smooth(aes(x=year, y=midwifedensity), method="loess", se=T, size=1, color="green") +
  geom_point(aes(x=year, y=mddensity, color= "Medical Doctor Density"), size = 0.8, alpha=1/5) +
  geom_smooth(aes(x=year, y=mddensity), method="loess", se=T, size=1, color="red") +
  labs(title = "Factors that may Influence the Mortality Rate in Mozambique",
                    y =" Variable Density (per 1000 live births)", 
                    x = "Years")+
  theme(legend.position = "bottom",
  plot.caption = element_text(hjust = 0), 
                    plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(breaks = c(2000,2002,2004,2006,2008,2010))  + theme(legend.title=element_blank())

scat44  
```
```

